Spectral partitioning identifies individual heterogeneity in the functional network topography of ventral and anterior medial prefrontal

### 2019-09-14 Manuscript.Rmd
discussion ready

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a diverse set of cognitive processes [@Hiser2018; @Kragel2018], and contain subregions that are part of the brain's default network [DN; @Buckner2008]. DN regions are characterized by a decrease in BOLD activity during externally oriented tasks  that require attention or cognitive control, in comparison with less-demanding task conditions or periods of rest [@Buckner2008; @Laird2009; @Mckiernan2003]. The DN can also be identified on the basis of a distinctive pattern of inter-region correlations in resting-state fMRI data [@Buckner2008; @Fox2005; @Greicius2003; @Yeo2011].

Many different cognitive task manipulations evoke patterns of brain activity that overlap with DN regions in ventral/anterior mPFC and in PCC. Examples include manipulations of self-referential thinking [@Gusnard2001; @Mitchell2005], memory [@Euston2012; @Schacter2007], affective regulation [@Reddan2018; @Schiller2008], and subjective valuation [@Bartra2013; @Clithero2014; @Kable2007; @Levy2011). Some of these task-related effects are thought to reflect processes integral to the functional role of the DN, such as internally oriented cognition, scene construction, and self-projection [@Hassabis2007]. For other task-related effects, such as subjective valuation (i.e., greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued), the degree of overlap with DN regions is only partial and the reason for the overlap is less obvious [@Acikalin2017]. Insofar as the DN shares a subset of nodes in common with the distributed brain system that supports valuation and other functions, this has potential to inform our theoretical understanding of the cognitive operations involved in those functions [@Northoff2012]. As a result, there is widespread interest in understanding the degree to which DN regions overlap topographically with task-related effects [@Dinicola2019; @Spreng2012].

@@

A considerable amount of meta-analytic work has been dedicated to characterizing the brain activity patterns associated with psychological processes in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across domains [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, activation patterns associated with subjective valuation and with the default network (DN) have been suggested to be inseparable in this area, with overlap also partially extending to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. This apparent neural overlap has important implications, as it has motivated theoretical proposals about ways in which these superficially dissimilar domains might involve a shared set of core cognitive processes [@Acikalin2017; @Clithero2014; @Northoff2012].

However, the interpretation of overlap in group-level data depends on the degree to which functional organization is heterogeneous across individuals. Recent studies have shown that heteromodal brain regions have considerable variability in functional connectivity across individuals [@Mueller2013], individual-specific functional topography can be occluded in aggregative estimations [@Braga2017; @Gordon2017; @Michalka2015; @Tobyne2018], and overlap in functional activation can vanish with increases in spatial precision [@Woo2014]. These findings suggest that group-level and meta-analysis-level overlap does not necessarily imply overlap in individual brains. However, our understanding of the individual-level heterogeneity in the functional topography of mPFC has been mostly descriptive so far [@Braga2017; @Braga2019; @Gordon2017]. To affirm that DN, valuation, and other functions share neural substrates in this area requires a method to reliably and precisely capture the functional topography of mPFC in isolated individuals, as well as a quantitative estimate of the degree of topographic heterogeneity across a large group of individuals.

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For 100 individuals, we subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use parcels that are several orders of magnitude larger than vertices as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, including identifying communities deterministically, constraining communities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values that relate to the strength of a node's community affiliation, and the ability to diagnose the connectedness of a network through examination of its resulting eigenvalues [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities [@Garcia2017], plus a silhouette analysis, suggested the binary partitioning was appropriate.

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. The precise spatial positioning of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individual. The idiosyncrasy in functional topography was particularly pronounced in mPFC, and was identified in both run-based and day-based analyses. Individual-specificity could theoretically arise from a variety of sources. For example, Individual variability could be due to shifts in functional organization that are independent of structural features [@Haxby2013; @Nenning2017], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. Alternatively, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to vary systematically across individuals [@Mackey2014]. Our results support this idea, echoing previous findings that DN is contained within sulci (in particular the superior rostral sulcus) in vmPFC [@Lopez2019]. This relationship was heterogeneous throughout the rest of the search space. DN was likely located in gyri in more dorsal mPFC regions, whereas this association differed across hemispheres in PCC. An important goal for future work will be to assess whether the network layout in these regions can also be predicted on the basis of other aspects of brain structure, such as myeloarchitecture [@Glasser2016] or structural connectivity [@Osher2016; @Saygin2011; @Saygin2016].

Network-partitioning methods such as SP are data-driven, and therefore provide no labeling information about the resulting communities. We circumvented this issue by independently identifying the DN community based on its coverage of area 7m, a region in PCC that was preferentially associated with the DN relative to subjective valuation in our meta-analysis. Our focus on this contrast was appropriate because of the notable topographical overlap between these processes in mPFC and PCC, despite being segregable elsewhere in the brain [@Acikalin2017]. We were able to apply labels derived from this group-level approach on the basis of the topography in PCC, where functional organization was more consistent across individuals. Because each community spanned both mPFC and PCC, the labels extended to mPFC where topography was more heterogeneous.

Our results extend previous work that described individual-specific brain organization. Several recent investigations have identified topographic heterogeneity using a different data aspect ratio than we used here [a small number of individuals and a large number of scanning sessions per individual; @Braga2017; @Braga2019; @Gordon2017]. Previous work has also shown that functional correlations among pre-defined cortical parcels are highly stable within an individual [@Gratton2018b; @Kong2018]. Here we were able to quantify the variability and stability of functional topography in a large sample at a fine, vertex-level spatial granularity, using moderately low amounts of data (down to a single 14 minute scan, although estimates become considerably more unreliable). The motivation to sub-divide DN also stems from recent work by Kernbach et al. [-@Kernbach2018], which identified specialized communication of parcels within DN with the rest of the brain in a large pool of individuals.

In addition to the technical advantages noted above, the SP algorithm offers analytical advantages specific to neuroscience. We found that SP outperformed a traditional seed-based correlation approach in capturing idiosyncratic functional topography. Community detection methods such as SP are stabilized by relying on  all pairwise correlations among cortical vertices (rather than correlations with an individual seed). In addition, we found we could increase the temporal stability of SP results by thresholding the underlying Fiedler vector. The magnitude of Fiedler vector values has been recently used to characterize the continuous connectivity profile of the insula with the rest of the brain, challenging the notion of discrete parcellations in that region [@Tian2018]. The combination of discrete classification and graded information yielded by SP provides additional flexibility and richness relative to some other clustering algorithms.

The community organization of PCC and mPFC was congruent with DN sub-networks A and B proposed by Braga & Buckner [-@Braga2017; @Braga2019]. The topography of our thresholded DN community closely matched network A, whereas our non-DN community included cortical territory that was not part of either DN network. Subthreshold vertices from the SP communities in turn overlapped with DN-B vertices. Our findings therefore complement the initial identification of DN sub-networks by quantifying the systematic variability of their underlying topography in a larger group of people. Understanding the interaction of networks A (DN), B, and non-DN is an important goal for future research. SP is related to methods that have gained traction recently for distinguishing functional cortical gradients [@Margulies2016; @Huntenburg2018]. Dedicated work could determine whether DN-A and DN-B form part of a gradual information processing sequence, or if their functions can be discretized. Regardless, this set of results collectively suggest that canonical DN regions can be topographically partitioned into DN and non-DN communities, and that the DN community can in turn be further divided into sub-networks A and B.

@@

Our findings show that the functional topography of mPFC is variable across a large pool of individuals, and that the SP algorithm is a useful tool for identifying individualized topography in a data-driven way. The ability to capture an individual's functional topography without the need of group priors is clinically relevant, as it could help characterize heterogenous changes in mPFC activity in disorders such as depression and schizophrenia [@Hiser2018]. It will be beneficial for future task-based fMRI experiments to be able to characterize where task-evoked activity is situated relative to an individual's overall mPFC organization. Our work was originally motivated by the apparent spatial overlap between subjective valuation and the default network [@Acikalin2017], but additional distinctions of interest include valuation versus episodic memory functions [@Euston2012], self-referential thought and theory of mind [@Mitchell2005], which have been grouped under the DN umbrella based on group averages [@Spreng2012], and potential topographic distinctions among multiple forms of valuation [@Clithero2014; @Shenhav2019; @Smith2010]. An individualized frame of reference will enhance the ability of future studies to gauge similarities and differences among brain activity patterns associated with diverse psychological domains.

>>>>

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a diverse set of cognitive processes [@Hiser2018; @Kragel2018], and contain subregions that are part of the brain's default network [DN; @Buckner2008; @Buckner2019]. DN regions are characterized by a decrease in BOLD activity during externally oriented tasks  that require attention or cognitive control, in comparison with less-demanding task conditions or periods of rest [@Buckner2008; @Laird2009; @Mckiernan2003]. The DN can also be identified on the basis of a distinctive pattern of inter-region correlations in resting-state fMRI data [@Buckner2008; @Fox2005; @Greicius2003; @Yeo2011].

Many different cognitive task manipulations evoke patterns of brain activity that overlap with DN regions in ventral/anterior mPFC and in PCC. Examples include manipulations of self-referential thinking [@Gusnard2001; @Mitchell2005], memory [@Euston2012; @Schacter2007], affective regulation [@Reddan2018; @Schiller2008], and subjective valuation [@Bartra2013; @Clithero2014; @Kable2007; @Levy2011]. Some of these task-related effects are thought to reflect processes integral to the functional role of the DN, such as internally oriented cognition, scene construction, and self-projection [@Hassabis2007]. For other task-related effects, such as subjective valuation (i.e., greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued), the degree of overlap with DN regions is only partial and the reason for the overlap is less obvious [@Acikalin2017]. Insofar as the DN shares a subset of nodes in common with the distributed brain system that supports valuation and other functions, this has potential to inform our theoretical understanding of the cognitive operations involved in those functions [@Northoff2012]. As a result, there is widespread interest in understanding the degree to which DN regions overlap topographically with task-related effects [@Buckner2019; @Dinicola2019; @Spreng2012].

@@

A considerable amount of meta-analytic work has been dedicated to characterizing the brain activity patterns associated with psychological processes in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across domains [@DelaVega2016; @Kragel2018; @Hiser2018]. For example, activation patterns associated with subjective valuation and with the default network (DN) have been suggested to be indistinguishable in mPFC, with overlap also partially extending to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. This apparent overlap of task-related effects with DN regions has important implications, as it has motivated theoretical proposals about ways in which these superficially dissimilar domains might involve a shared set of core cognitive processes [@Acikalin2017; @Clithero2014; @Northoff2012].

However, the interpretation of overlap in group-level data depends on the degree to which functional organization is heterogeneous across individuals. Recent studies have shown that heteromodal brain regions have considerable variability in functional connectivity across individuals [@Mueller2013], individual-specific functional topography can be occluded in aggregative estimations [@Braga2017; @Gordon2017; @Michalka2015; @Tobyne2018], and overlap in functional activation can vanish with increases in spatial precision [@Woo2014]. These findings suggest that group-level and meta-analysis-level overlap does not necessarily imply overlap in individual brains. To date, our understanding of the individual-level heterogeneity in the functional topography of mPFC has been mostly descriptive [@Braga2017; @Braga2019; @Gordon2017]. A strong test of the overlap between task-related effects and DN regions would require a method to reliably and precisely capture the functional topography of mPFC in isolated individuals, as well as a quantitative estimate of the degree of topographic heterogeneity across a large group of individuals.

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For each of 100 individuals, we subdivided canonical DN regions into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use parcels that are several orders of magnitude larger than vertices as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, including identifying communities deterministically, constraining communities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values that relate to the strength of a node's community affiliation, and the ability to diagnose the connectedness of a network through examination of its resulting eigenvalues [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities [@Garcia2017], as well as a silhouette analysis, suggested the binary partitioning was appropriate.

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral and anterior/superior mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. The precise spatial positioning of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individual. The idiosyncrasy in functional topography was particularly pronounced in mPFC, and was identified in both run-based and day-based analyses. Individual-specificity could theoretically arise from a variety of sources. For example, individual variability could be due to shifts in functional organization that are independent of structural features [@Haxby2013; @Nenning2017], and could relate to the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. Alternatively, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to vary systematically across individuals [@Mackey2014]. Our results offer some support for this idea, echoing previous findings that DN is contained within sulci (in particular the superior rostral sulcus) in vmPFC [@Lopez2019]. Structure/function associations were heterogeneous in other regions; DN tended to be located in gyri in more superior mPFC regions, whereas the association differed across hemispheres in PCC. Future studies should further characterize this heterogeneous relationship. Another important goal for future work will be to assess whether the network layout in these regions can also be predicted on the basis of other aspects of brain structure, such as myeloarchitecture [@Glasser2016] or structural connectivity [@Osher2016; @Saygin2011; @Saygin2016].

Network-partitioning methods such as SP are data-driven, and therefore provide no labeling information about the resulting communities. We circumvented this issue by independently identifying the DN community based on its coverage of area 7m, a region in PCC that was preferentially associated with the DN relative to subjective valuation in our meta-analysis. We were able to apply labels derived from this group-level approach on the basis of the topography in PCC, where functional organization was more consistent across individuals. Because each community spanned both mPFC and PCC, the labels extended to mPFC where topography was more heterogeneous.

Our results extend previous work that described individual-specific brain organization. Several recent investigations have identified topographic heterogeneity using a different data aspect ratio than we used here [a small number of individuals and a large number of scanning sessions per individual; @Braga2017; @Braga2019; @Gordon2017]. Previous work has also shown that functional correlations among pre-defined cortical parcels are highly stable within an individual [@Gratton2018b; @Kong2018]. Here we were able to quantify the variability and stability of functional topography in a large sample at a fine, vertex-level spatial granularity, using moderately low amounts of data (down to a single 14 minute scan, although estimates based on more data were more reliable). The motivation to subdivide DN also stems from recent work by Kernbach et al. [-@Kernbach2018], which identified specialized communication of parcels within DN with the rest of the brain in a large pool of individuals.

In addition to the technical advantages noted above, the SP algorithm offers analytical advantages specific to neuroscience. We found that SP outperformed a traditional seed-based correlation approach in capturing idiosyncratic functional topography. Community detection methods such as SP are stabilized by relying on  all pairwise correlations among cortical vertices (rather than correlations with an individual seed). In addition, we found we could threshold the underlying Fiedler vector on the basis of the temporal stability of SP results. The magnitude of Fiedler vector values has been recently used to characterize the continuous connectivity profile of the insula with the rest of the brain, challenging the notion of discrete parcellations in that region [@Tian2018]. The combination of discrete classification and graded information yielded by SP provides additional flexibility and richness relative to some other clustering algorithms.

The community organization of PCC and mPFC was congruent with DN sub-networks A and B proposed by Braga & Buckner [-@Braga2017; @Braga2019]. The topography of our thresholded DN community closely matched network A, whereas our non-DN community included cortical territory that was not part of either DN network. Subthreshold vertices from the SP communities in turn overlapped with DN-B vertices. Our findings therefore complement the initial identification of DN sub-networks by quantifying the systematic variability of their underlying topography in a larger group of people. Understanding the interaction of networks DN-A, DN-B, and non-DN is an important goal for future research. SP is related to methods that have gained traction recently for distinguishing functional cortical gradients [@Margulies2016; @Huntenburg2018]. A valuable goal for future work would be to asses whether DN-A and DN-B form part of a gradual information processing sequence, or if their functions can be discretized. Regardless, this set of results collectively suggested that canonical DN regions can be topographically partitioned into DN and non-DN communities, and that the DN community can in turn be further divided into sub-networks A and B.

@@

Our findings show that the functional topography of mPFC is variable across a large pool of individuals, and that the SP algorithm is a useful tool for identifying individualized topography in a data-driven way. The ability to capture an individual's functional topography without the need for group priors is clinically relevant, as it could help characterize heterogenous changes in mPFC activity in disorders such as depression and schizophrenia [@Hiser2018]. It will be beneficial for future task-based fMRI experiments to be able to characterize where task-evoked activity is situated relative to an individual's overall mPFC organization. Our work is relevant to interpreting the overlap of DN regions with task-related brain activity in numerous cognitive domains, including valuation [@Acikalin2017; @Shenhav2019], memory [@Euston2012], and self-referential thought [@Mitchell2005]. An individualized frame of reference will enhance the ability of future studies to gauge similarities and differences among brain activity patterns associated with diverse psychological domains.
